### 2023-06-24 manuscript/results.tex
finished results

\section{Results}
We implemented the prefix-free graphs as a Rust package.
This package contains two binary crates and one library crate.
Binary crates are executables that serve the purpose of creating prefix-free graphs from FASTA and GFA formats.
The binary crates are named \texttt{fasta2pfg} and \texttt{gfa2pfg}, and their usage is as follows:

\begin{minted}{bash}
fasta2pfg -t triggers.txt < pangenome.fna > pfg.gfa

gfa2pfg -t triggers.txt < pangenome.gfa > pfg.gfa
\end{minted}

We compared the running time of the construction algorithm with several pangenome construction tools, namely with the PanGenome Graph Builder \cite{garrison2023building}, VG \cite{garrison2018variation}, and Minigraph-Cactus \cite{hickey2023pangenome}.
For comparison, we used up to 256 SARS-CoV-2 sequences.
These sequences have lengths of around 29.9 kbp and high nucleotide similarity.
The resulting running times are shown in Table \ref{tab:comparison}.

\input{comparison_table.tex}

The library crate provides an interface for working with prefix-free graphs, mainly an iterator of a suffix array in sublinear space and linear time.
The library can be used from within the Rust programming language as follows:

\begin{minted}{rust}
let pfg = PFG::load("pfg.gfa");

for (i, (sa_i, id_i, pos_i)) in pfg.iter().enumerate() {
    println!("{}\t{}\t{}\t{}", i, sa_i, id_i, pos_i);
}
\end{minted}

In addition to the suffix array value, the iterator provides the segment ID and position values.
These are useful for the computation of several related data structures.
Similarly to \citet{2019boucher}, we can use them to compute the Burrows-Wheeler transform $\texttt{BWT}[i]$ by storing the preceding characters of each segment and then reporting the character at position $\texttt{pos}[i]-1$ of a segment $\texttt{id}[i]$.
This computation allows for space-efficient construction of r-index \cite{gagie2020fully} similar to MONI implementation \cite{rossi2022moni}, space-efficient construction of Wheeler graphs similar in nature to the implementation in \citet{2022pfwg}, and with the use of predecessor queries, space-efficient construction of a tag array introduced in \citet{2022maria}.