### 2023-08-09 manuscript/content/table_segment.tex

\begin{table}
    \caption{
    Segment table containing an id, length, starting positions in the concatenated pangenome and ranks of the right contexts of these positions for each segment. \label{tab:segment}}
    \begin{tabular}{rrll}
        \toprule
        id & length & starts & right context ranks \\
        \midrule
        0 & 4 & 9           & 9         \\
        1 & 3 & 15, 1       & 13, 14    \\
        2 & 5 & 18, 5, 11   & 1, 2, 3   \\
        3 & 3 & 8, 14, 0    & 4, 5, 6   \\
        4 & 4 & 16          & 7         \\
        5 & 5 & 2           & 8         \\
        \bottomrule
    \end{tabular}
\end{table}