### 2023-08-09 manuscript/content/table_suffix.tex

\begin{table*}
    \caption{
    Suffix table with the suffix array \texttt{SA}, longest common prefix array \texttt{LCP}, segment ID array \texttt{ID} and segment positions array \texttt{pos} of the running example. Suffixes of segment join from position \texttt{SA}[i] added for easier interpretation. Darker colours represent the suffix of the current segment. \label{tab:suffix}
    }
    \begin{tabular}{rrrrrl}
        \toprule
        i & \texttt{SA}[i] & \texttt{LCP}[i] & \texttt{ID}[i] & \texttt{pos}[i] & segment join[\texttt{SA}[i]..] \\
        \midrule
         0  & 30 & -1 & 6 & 0 & \texttt{\textcolor{gray}{\$}}                                     \\
         1  & 29 &  0 & 5 & 5 & \texttt{\textcolor{gray}{\#\$}}                                   \\
         2  &  4 &  1 & 0 & 4 & \texttt{\textcolor{gray}{\#ACG\#ACT..\#CAC\#CGAC\#CGTAC\#\$}}     \\
         3  &  8 &  3 & 1 & 3 & \texttt{\textcolor{gray}{\#ACT..\#CAC\#CGAC\#CGTAC\#\$}}          \\
         4  & 14 &  1 & 2 & 5 & \texttt{\textcolor{gray}{\#CAC\#CGAC\#CGTAC\#\$}}                 \\
         5  & 18 &  2 & 3 & 3 & \texttt{\textcolor{gray}{\#CGAC\#CGTAC\#\$}}                      \\
         6  & 23 &  3 & 4 & 4 & \texttt{\textcolor{gray}{\#CGTAC\#\$}}                            \\
         7  & 13 &  0 & 2 & 4 & \texttt{.\textcolor{gray}{\#CAC\#CGAC\#CGTAC\#\$}}                \\
         8  & 12 &  1 & 2 & 3 & \texttt{..\textcolor{gray}{\#CAC\#CGAC\#CGTAC\#\$}}               \\
         9  & 27 &  0 & 5 & 3 & \texttt{AC\textcolor{gray}{\#\$}}                                 \\
        10  &  2 &  3 & 0 & 2 & \texttt{AC\textcolor{gray}{\#ACG\#ACT..\#CAC\#CGAC\#CGTAC\#\$}}   \\
        11  & 16 &  3 & 3 & 1 & \texttt{AC\textcolor{gray}{\#CGAC\#CGTAC\#\$}}                    \\
        12  & 21 &  5 & 4 & 2 & \texttt{AC\textcolor{gray}{\#CGTAC\#\$}}                          \\
        13  &  0 &  2 & 0 & 0 & \texttt{ACAC\textcolor{gray}{\#ACG\#ACT..\#CAC\#CGAC\#CGTAC\#\$}} \\
        14  &  5 &  2 & 1 & 0 & \texttt{ACG\textcolor{gray}{\#ACT..\#CAC\#CGAC\#CGTAC\#\$}}       \\
        15  &  9 &  2 & 2 & 0 & \texttt{ACT..\textcolor{gray}{\#CAC\#CGAC\#CGTAC\#\$}}            \\
        16  & 28 &  0 & 5 & 4 & \texttt{C\textcolor{gray}{\#\$}}                                  \\
        17  &  3 &  2 & 0 & 3 & \texttt{C\textcolor{gray}{\#ACG\#ACT..\#CAC\#CGAC\#CGTAC\#\$}}    \\
        18  & 17 &  2 & 3 & 2 & \texttt{C\textcolor{gray}{\#CGAC\#CGTAC\#\$}}                     \\
        19  & 22 &  4 & 4 & 3 & \texttt{C\textcolor{gray}{\#CGTAC\#\$}}                           \\
        20  &  1 &  1 & 0 & 1 & \texttt{CAC\textcolor{gray}{\#ACG\#ACT..\#CAC\#CGAC\#CGTAC\#\$}}  \\
        21  & 15 &  4 & 3 & 0 & \texttt{CAC\textcolor{gray}{\#CGAC\#CGTAC\#\$}}                   \\
        22  &  6 &  1 & 1 & 1 & \texttt{CG\textcolor{gray}{\#ACT..\#CAC\#CGAC\#CGTAC\#\$}}        \\
        23  & 19 &  2 & 4 & 0 & \texttt{CGAC\textcolor{gray}{\#CGTAC\#\$}}                        \\
        24  & 24 &  2 & 5 & 0 & \texttt{CGTAC\textcolor{gray}{\#\$}}                              \\
        25  & 10 &  1 & 2 & 1 & \texttt{CT..\textcolor{gray}{\#CAC\#CGAC\#CGTAC\#\$}}             \\
        26  &  7 &  0 & 1 & 2 & \texttt{G\textcolor{gray}{\#ACT..\#CAC\#CGAC\#CGTAC\#\$}}         \\
        27  & 20 &  1 & 4 & 1 & \texttt{GAC\textcolor{gray}{\#CGTAC\#\$}}                         \\
        28  & 25 &  1 & 5 & 1 & \texttt{GTAC\textcolor{gray}{\#\$}}                               \\
        29  & 11 &  0 & 2 & 2 & \texttt{T..\textcolor{gray}{\#CAC\#CGAC\#CGTAC\#\$}}              \\
        30  & 26 &  1 & 5 & 2 & \texttt{TAC\textcolor{gray}{\#\$}}                                \\
        \bottomrule
    \end{tabular}
\end{table*}